Revisited hidden weight bit function

### 2024-11-04 RevisedHiddenWeight.tex
Clarifications and typo corrections in Thm. 3 proof

\begin{theorem}\label{theorem:bound_D_nka_generalized}
    Let $t, n\in \N$ with $n$ even and $t\le n/2$. Define $\lambda:=t/n\in\left[0,\frac 12\right]$, as well as:
    \[
        \mu:=\mu(\lambda):=
        \begin{cases}
            \frac{\lambda+1}{2}+\frac12\log_2\left(\frac{\left(-\lambda^2+2\lambda+\lambda\sqrt{\lambda^2-4\lambda+2}\right)^\lambda}{\left(1-\lambda+\sqrt{\lambda^2-4\lambda+2}\right)^{2\lambda-1}}\right)&\text{if $\lambda>\frac 16$,}\\
            1-\lambda&\text{if $\lambda\leq\frac 16$.}\\
        \end{cases}
    \]
    Then for all $a\in\F_2^n$ and $k\in[0,n]$, the following hold true:
    \[
        |\wtk{d_{t,n}}{k}(a)|\leq 2^{\mu n},\quad |\wt{d_{t,n}}(a)|\leq 1+n\cdot 2^{\mu n}.
    \]
\end{theorem}

\begin{proof}
    To begin, observe that the inequality $|\wt{d_{t,n}}(a)|\leq 1+n\cdot 2^{\mu n}$ will follow from the inequality $|\wtk{d_{t,n}}{k}(a)|\leq 2^{\mu n}$ using precisely the arguments of the proof of Corollary~\ref{cor:bound_walsh_cauchy}, so that we are only required to prove the first inequality.

@@

   This leaves us with the following cases to consider for $(p,q,r;u,v)$:
    \begin{gather*}
        (t,0,0;n-2t,0),\quad(0,t,0;n-2t,0),\quad(0,0,t;n-2t,0),\\
         (t,0,0;0,n-2t),\quad(0,t,0;0,n-2t),\quad(0,0,t;0,n-2t).
    \end{gather*}
    Since swapping both $(p,q)$ and $(u,v)$ has the same effect as replacing $z$ by $-z$ in $P_{a,t}(z)$ and therefore does not change the maximum of $|P_{a,t}(z)|$ on $\dis$, we can further restrict ourselves to the following tuples for $(p,q,r;u,v)$:
    \[
        (t,0,0;n-2t,0),\quad(0,t,0;n-2t,0),\quad(0,0,t;n-2t,0).
    \]
    Let us start by bounding $|P_{a,t}(z)|$ on $\dis$ in case $(p,q,r;u,v)=(t,0,0;n-2t,0)$. Since $t=\lambda n$, we can write:
    \[
        P_{a,t}(z)=\left((-z^2+2z+1)^{\lambda}(1+z)^{1-2\lambda}\right)^n.
    \]
    Writing $z=\ee^{\ii\alpha}$ and $z^2=\ee^{2\ii\alpha}$ for $\alpha\in[-\pi,\pi]$, and using that $|w|^2=w\overline w$ for every complex number $w$, we get:
    \begin{align*}
        |-z^2+2z+1|^2&=\left(-\ee^{2\ii\alpha}+2\ee^{\ii\alpha}+1\right)\left(-\ee^{-2\ii\alpha}+2\ee^{-\ii\alpha}+1\right)\\
        &=6-\ee^{2\ii\alpha}-\ee^{-2\ii\alpha}\\
        &=6-2\cos (2\alpha),
    \end{align*}
    where the last equality uses that $\ee^{2\ii\alpha}=\cos(2\alpha)+\ii\sin(2\alpha)$. We similarly obtain that $|1+z|^2=2+2\cos\alpha$. Combining both, we get:   
    \[
        |P_{a,t}(z)|^2=\left(\left|-z^2+2z+1\right|^{2\lambda}|1+z|^{2(1-2\lambda)}\right)^n=\left((6-2\cos(2\alpha))^\lambda(2+2\cos\alpha)^{1-2\lambda}\right)^n.
    \]
    The goal will be to prove that the maximum of $g(\alpha):=(6-2\cos(2\alpha))^\lambda(2+2\cos\alpha)^{1-2\lambda}$ for $\alpha\in[-\pi,\pi]$ is equal to $2^{2\mu}$; the maximum of $|P_{a,t}(z)|$ on $\dis$ will then be $\left(2^{2\mu}\right)^{n/2}=2^{\mu n}$. Since $g$ is an even function, it is enough to focus on the interval $\alpha\in[0,\pi]$.

    To begin, computing the first derivative of $g$ and applying the \textit{double angle formulas} $\cos(2\alpha)=2\cos^2\alpha-1$ and $\sin(2\alpha)=2\cos\alpha\sin\alpha$ gives:
    \[
        \frac{\dd}{\dd\alpha}g(\alpha)=8(6-2\cos(2\alpha))^{\lambda-1}(2+2\cos\alpha)^{-2\lambda}\sin\alpha\left(\cos^2\alpha+2\lambda\cos\alpha+4\lambda-2\right).
    \]
    The only interesting zeros of $\frac{\dd}{\dd\alpha}g(\alpha)$ are those of the last factor 
    $\cos^2\alpha+2\lambda\cos\alpha+4\lambda-2$: those of the factor $8(6-2\cos(2\alpha))^{\lambda-1}(2+2\cos\alpha)^{-2\lambda}$ will make $g(\alpha)$ vanish and will therefore not maximize $g(\alpha)$, and those for $\sin\alpha$ are $\alpha=0$ and $\alpha=\pi$.

    Our first claim is that the equation $\mathcal E:\cos^2\alpha+2\lambda\cos\alpha+4\lambda-2=0$ either has a unique solution $\cos\alpha=-\lambda+\sqrt{\lambda^2-4\lambda+2}$ if $\lambda\geq\frac 16$, and has no solution $\cos\alpha$ otherwise. To see this, we consider $\mathcal E$ as an equation of second degree in $\cos\alpha$, it leads to the two solutions $\cos\alpha=-\lambda\pm\sqrt{\lambda^2-4\lambda+2}$. We can exclude the solution $\cos\alpha=-\lambda-\sqrt{\lambda^2-4\lambda+2}$ since $\cos\alpha\in[-1,1]$. The other solution needs to be excluded for the same reason if $\lambda<\frac 16$.

    Our next claim is that $\frac{\dd^2}{\dd^2\alpha}g(\alpha)\vert_{\alpha=0}$ has the same sign as $6\lambda -1$. Indeed, by definition of the second derivative, we have:
    \begin{align*}
        \frac{\dd^2}{\dd^2\alpha}g(\alpha)\vert_{\alpha=0}
        &=\lim_{\alpha\to 0}\frac{\frac{\dd}{\dd\alpha}g(\alpha)}{\alpha}\\
        &=\lim_{\alpha\to 0}8(6-2\cos(2\alpha))^{\lambda-1}(2+2\cos\alpha)^{-2\lambda}\cdot\frac{\sin\alpha}{\alpha}\cdot\left(\cos^2\alpha+2\lambda\cos\alpha+4\lambda-2\right)\\
        &=8\cdot4^{\lambda-1}\cdot 4^{-2\lambda}\cdot 1\cdot(6\lambda -1)\\
        &=2^{1-2\lambda}(6\lambda -1).
    \end{align*}

    Therefore, in case $\lambda<\frac 16$, we can conclude from the two claims that $g(\alpha)$ reaches its maximum at $\alpha_0=0$, and this maximum is then equal to $2^{2-2\lambda}=2^{2\mu}$, as required. For $\lambda=\frac 16$, the solution to the equation $\mathcal E$ is $\cos\alpha=1$, again implying $\alpha_0=0$, and we obtain the same maximum of $2^{2\mu}$. In case $\lambda>\frac 16$, the two claims imply that $g(\alpha)$ reaches its maximum at the two values $\alpha_0\in[0,pi]$ for which $\cos\alpha_0=-\lambda+\sqrt{\lambda^2-4\lambda+2}$. For such $\alpha$, the \textit{double angle formula} for the cosine implies that $\cos(2\alpha_0)=2\lambda^2-4\lambda+2-2\lambda\sqrt{\lambda^2-4\lambda+2}$. If we replace $\cos(\alpha_0)$ and $\cos(2\alpha_0)$ by their respective values, we obtain that the maximum of $g(\alpha)$ on $\alpha\in[0,\pi]$ is given by the following value:
    \[
        g(\alpha_0)=\frac{\left(-8\lambda^2+16\lambda+8\lambda\sqrt{\lambda^2-4\lambda+2}\right)^{\lambda}}{\left(2-2\lambda+2\sqrt{\lambda^2-4\lambda+2}\right)^{2\lambda-1}}=2^{2\mu}.
    \]
    This is exactly what we wanted to prove, and we have therefore covered the case $(p,q,r;u,v)=(t,0,0;n-2t,0)$ entirely. In a similar fashion, it can be proven that $|P_{a,t}(z)|$ is also bounded by $2^{\mu n}$ on $\dis$  for the remaining two cases of $(p,q,r;u,v)$.
\end{proof}

The curve $\lambda\mapsto\mu(\lambda)$ defined in the above Theorem~\ref{theorem:bound_D_nka_generalized} is represented in Figure~\ref{fig:curve_exponents_lambda_mu}.

\begin{figure}
	\centering
	\includegraphics[width=9cm]{curve_exponents_lambda_mu.pdf}
	\caption{The curve $\lambda\mapsto\mu(\lambda)$ from Theorem~\ref{theorem:bound_D_nka_generalized}}
	\label{fig:curve_exponents_lambda_mu}
\end{figure}

>>>>

\begin{theorem}\label{theorem:bound_D_nka_generalized}
    Let $t, n\in \N$ with $n$ even and $t\le n/2$. Define $\lambda:=t/n\in\left[0,\frac 12\right]$, as well as:
    \[
        \mu:=\mu(\lambda):=
        \begin{cases}
            \frac{\lambda+1}{2}+\frac12\log_2\left(\frac{\left(-\lambda^2+2\lambda+\lambda\sqrt{\lambda^2-4\lambda+2}\right)^\lambda}{\left(1-\lambda+\sqrt{\lambda^2-4\lambda+2}\right)^{2\lambda-1}}\right)&\text{if $\lambda>\frac 16$,}\\
            1-\lambda&\text{if $\lambda\leq\frac 16$.}\\
        \end{cases}
    \]
    Then for all $a\in\F_2^n$ and $k\in[1,n]$, the following hold true:
    \[
        |\wtk{d_{t,n}}{k}(a)|\leq 2^{\mu n},\quad |\wt{d_{t,n}}(a)|\leq 1+n\cdot 2^{\mu n}.
    \]
\end{theorem}

\begin{proof}
    To begin, observe that the inequality $|\wt{d_{t,n}}(a)|\leq 1+n\cdot 2^{\mu n}$ will follow from the inequality $|\wtk{d_{t,n}}{k}(a)|\leq 2^{\mu n}$ using precisely the arguments of the proof of Corollary~\ref{cor:bound_walsh_cauchy}, so that we are only required to prove the first inequality. Also, since the case $\lambda=\frac 12$ corresponds to Theorem~\ref{theorem:bound_D_nka}, we will henceforth assume $\lambda<\frac 12$.

@@

   This leaves us with the following cases to consider for $(p,q,r;u,v)$:
    \begin{gather*}
        (t,0,0;n-2t,0),\quad(0,t,0;n-2t,0),\quad(0,0,t;n-2t,0),\\
         (t,0,0;0,n-2t),\quad(0,t,0;0,n-2t),\quad(0,0,t;0,n-2t).
    \end{gather*}
    Since swapping both $(p,q)$ and $(u,v)$ has the same effect as replacing $z$ by $-z$ in $P_{a,t}(z)$ and therefore does not change the maximum of $|P_{a,t}(z)|$ on $\dis$, we can further restrict ourselves to the following tuples for $(p,q,r;u,v)$:
    \[
        (t,0,0;n-2t,0),\quad(0,t,0;n-2t,0),\quad(0,0,t;n-2t,0).
    \]
    Let us start by bounding $|P_{a,t}(z)|$ on $\dis$ in case $(p,q,r;u,v)=(t,0,0;n-2t,0)$. Since $t=\lambda n$, we can write:
    \[
        P_{a,t}(z)=\left((-z^2+2z+1)^{\lambda}(1+z)^{1-2\lambda}\right)^n.
    \]
    Further writing $z=\ee^{\ii\alpha}$ and $z^2=\ee^{2\ii\alpha}$ for $\alpha\in[-\pi,\pi]$, and using that $|w|^2=w\overline w$ for every complex number $w$, we get:
    \begin{align*}
        |-z^2+2z+1|^2&=\left(-\ee^{2\ii\alpha}+2\ee^{\ii\alpha}+1\right)\left(-\ee^{-2\ii\alpha}+2\ee^{-\ii\alpha}+1\right)\\
        &=6-\ee^{2\ii\alpha}-\ee^{-2\ii\alpha}\\
        &=6-2\cos (2\alpha),
    \end{align*}
    where the last equality uses that $\ee^{2\ii\alpha}=\cos(2\alpha)+\ii\sin(2\alpha)$. We similarly obtain that $|1+z|^2=2+2\cos\alpha$. Combining both, we get:   
    \[
        |P_{a,t}(z)|^2=\left(\left|-z^2+2z+1\right|^{2\lambda}|1+z|^{2(1-2\lambda)}\right)^n=\left((6-2\cos(2\alpha))^\lambda(2+2\cos\alpha)^{1-2\lambda}\right)^n.
    \]
    The goal will be to prove that the maximum of $g(\alpha):=(6-2\cos(2\alpha))^\lambda(2+2\cos\alpha)^{1-2\lambda}$ for $\alpha\in[-\pi,\pi]$ is equal to $2^{2\mu}$; the maximum of $|P_{a,t}(z)|$ on $\dis$ will then be $\left(2^{2\mu}\right)^{n/2}=2^{\mu n}$. Since $g$ is an even function, it is enough to focus on the interval $\alpha\in[0,\pi]$.

    To begin, computing the first derivative of $g$ and applying the double angle formulas $\cos(2\alpha)=2\cos^2\alpha-1$ and $\sin(2\alpha)=2\cos\alpha\sin\alpha$ gives:
    \[
        \frac{\dd}{\dd\alpha}g(\alpha)=8(6-2\cos(2\alpha))^{\lambda-1}(2+2\cos\alpha)^{-2\lambda}\sin\alpha\left(\cos^2\alpha+2\lambda\cos\alpha+4\lambda-2\right).
    \]
    The only interesting zeros of $\frac{\dd}{\dd\alpha}g(\alpha)$ are those of the last factor 
    $\cos^2\alpha+2\lambda\cos\alpha+4\lambda-2$: those of the factor $8(6-2\cos(2\alpha))^{\lambda-1}(2+2\cos\alpha)^{-2\lambda}$ will make $g(\alpha)$ vanish and will therefore not maximize $g(\alpha)$, and those for $\sin\alpha$ are $\alpha=0$ and $\alpha=\pi$, with corresponding values $g(0)=2^{2-2\lambda}$ and $g(\pi)=0$.

    Our first claim is that the equation $\mathcal E:\cos^2\alpha+2\lambda\cos\alpha+4\lambda-2=0$ has a unique solution $\cos\alpha=-\lambda+\sqrt{\lambda^2-4\lambda+2}$ if $\lambda\geq\frac 16$ and has no solution $\cos\alpha$ otherwise. To see this, we consider $\mathcal E$ as an equation of second degree in $\cos\alpha$, which leads to the two solution candidates $\cos\alpha=-\lambda\pm\sqrt{\lambda^2-4\lambda+2}$. We can exclude the solution $\cos\alpha=-\lambda-\sqrt{\lambda^2-4\lambda+2}$ since $\cos\alpha\in[-1,1]$. The other solution needs to be excluded for the same reason if $\lambda<\frac 16$.

    Our next claim is that $\frac{\dd^2}{\dd^2\alpha}g(\alpha)\vert_{\alpha=0}$ has the same sign as $6\lambda -1$. To see why, we apply the definition of the second derivative:
    \begin{align*}
        \frac{\dd^2}{\dd^2\alpha}g(\alpha)\vert_{\alpha=0}
        &=\lim_{\alpha\to 0}\frac{\frac{\dd}{\dd\alpha}g(\alpha)}{\alpha}\\
        &=\lim_{\alpha\to 0}8(6-2\cos(2\alpha))^{\lambda-1}(2+2\cos\alpha)^{-2\lambda}\cdot\frac{\sin\alpha}{\alpha}\cdot\left(\cos^2\alpha+2\lambda\cos\alpha+4\lambda-2\right)\\
        &=8\cdot4^{\lambda-1}\cdot 4^{-2\lambda}\cdot 1\cdot(6\lambda -1)\\
        &=2^{1-2\lambda}(6\lambda -1).
    \end{align*}

    Therefore, in case $\lambda<\frac 16$, we can conclude from the two claims that $g(\alpha)$ reaches its maximum at $\alpha_0=0$, and this maximum is then equal to $2^{2-2\lambda}=2^{2\mu}$, as required. For $\lambda=\frac 16$, the solution to the equation $\mathcal E$ is $\cos\alpha=1$, again implying that $g(\alpha)$ reaches its maximum at $\alpha_0=0$, and we obtain the same maximum of $2^{2\mu}$. In case $\lambda>\frac 16$, the two claims imply that $g(\alpha)$ reaches its maximum at the two values $\alpha_0\in[0,\pi]$ for which $\cos\alpha_0=-\lambda+\sqrt{\lambda^2-4\lambda+2}$. For such $\alpha$, the double angle formula for the cosine implies that $\cos(2\alpha_0)=2\lambda^2-4\lambda+2-2\lambda\sqrt{\lambda^2-4\lambda+2}$. If we replace $\cos\alpha_0$ and $\cos(2\alpha_0)$ by their respective values, we obtain that the maximum of $g(\alpha)$ for $\alpha\in[0,\pi]$ is given by the following value:
    \[
        g(\alpha_0)=\frac{\left(-8\lambda^2+16\lambda+8\lambda\sqrt{\lambda^2-4\lambda+2}\right)^{\lambda}}{\left(2-2\lambda+2\sqrt{\lambda^2-4\lambda+2}\right)^{2\lambda-1}}=2^{2\mu}.
    \]
    This is exactly what we wanted to prove, and we have therefore covered the case $(p,q,r;u,v)=(t,0,0;n-2t,0)$ entirely. In a similar fashion, it can be proven that $|P_{a,t}(z)|$ is also bounded by $2^{\mu n}$ on $\dis$ for the remaining two cases of $(p,q,r;u,v)$. In fact, for $(p,q,r;u,v)=(0,t,0;n-2t,0)$, it is enough to replace $g(\alpha)$ by $h_1(\alpha):=(6-2\cos(2\alpha))^\lambda(2-2\cos\alpha)^{1-2\lambda}$, which has the same maximum as $g(\alpha)$ because $h_1\left(\frac{\pi}{2}+\alpha\right)=g\left(\frac{\pi}{2}-\alpha\right)$; for $(p,q,r;u,v)=(0,0,t;n-2t,0)$, it is enough to replace $g(\alpha)$ by $h_2(\alpha):=(2+2\cos(2\alpha))^\lambda(2+2\cos\alpha)^{1-2\lambda}$, whose maximum is bounded by the maximum of $g(\alpha)$ because $h_2(\alpha)\leq g(\alpha)$.
\end{proof}

The curve of $\lambda\mapsto\mu(\lambda)$ defined in the above Theorem~\ref{theorem:bound_D_nka_generalized} is represented in Figure~\ref{fig:curve_exponents_lambda_mu}.

\begin{figure}
	\centering
	\includegraphics[width=10cm]{curve_exponents_lambda_mu.pdf}
	\caption{The curve of $\lambda\mapsto\mu(\lambda)$ with $0\leq\lambda\leq\frac 12$ from Theorem~\ref{theorem:bound_D_nka_generalized}}
	\label{fig:curve_exponents_lambda_mu}
\end{figure}
